### 2020-07-02 Thesis-Chapter1-Introduction.tex
Writing and editing comments

Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or metrics. Each attempt at a solution matters,
has unintended consequences, and often exists in a pattern of chronic policy failure.

@@

For these reasons, it is time that the technical community moves the debate forward.

@@

\principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
absolute privacy is not one of them. In some contexts, absolute privacy is proper, and in almost all contexts,
reasonable privacy should be presumed. However, privacy is not an absolute right, and investigators should have access
to some classes of data. Access to these classes of data is important today and will be increasingly important in the
future.}

\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
potential for abuse, international reception, economic impact, and need for transparency.}

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
right balance of risk given the threats under consideration--which includes regulatory action.}

>>>>

Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.

@@

For these reasons, it is important that the technical community keeps moving the debate forward.

@@

\principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
absolute privacy in all contexts is not one of them. In some contexts, absolute privacy is proper, and in all contexts,
reasonable privacy should be presumed. However, absolute privacy does not always apply, and investigators should have
access to some classes of data. Access to certain classes of digital data is important today and will be increasingly
important in the future.}

\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
potential for abuse, international consequences, economic impact, and need for transparency.}

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
right balance of risk given the threats under consideration--including the threat of regulatory action.}

### 2020-07-02 Thesis-Chapter2-Background.tex
Writing and editing comments

\Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
performed by using some \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. The cipher is the tool or
algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
cipher ensures that only those parties with the correct secret information, the \ac{key}, can perform encryption or
decryption on the text.

@@

There are two major cryptographic protocol families--symmetric and asymmetric. In traditional \ac{sym-crypto},
encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
securely. In \ac{asym-crypto}, also called \ac{public-key-crypto}, encryption and decryption are performed with two
paired keys, called the public key and the private key. The innovation of \ac{public-key-crypto} is that the public key
is not secret, and using this technique, two parties can communicate securely without requiring an a previously agreed
upon shared secret.

One can also define two major cryptographic applications--securing \acf{DIM} and \acf{DAR}. Each application is a
different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
perform authentication and establish short-lived \ac{sym-crypto} sessions keys; the sessions keys perform the actual
encryption of the data in transit. Important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}, which
ensures that a leaked private key or session key does not compromise any other private key or session key, and
\ac{replay-protection}, which ensures that messages cannot be replayed by an attacker without detection
\cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for \ac{DIM}.

\Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
encryption-time, as \ac{DIM} session keys are, these keys are either derived from a password stored in human memory or
are stored somewhere in computer memory, which often takes place with the assistance of dedicated hardware.
\Ac{disk-encryption} for laptops and mobile devices is an example of encryption for \ac{DAR}.

@@

Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
from the composition of their parts, and innumerable primitives and technologies rely on cryptography.

@@

Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
systems that directly rely on these technologies include the financial system, local and global; industrial control
systems and critical infrastructure; health information systems; private, government, and military communications; and
the integrity of the internet as a whole.

@@

The section provides an overview of encryption from ancient to contemporary times.

@@

Computerized encryption is a new technology, but the field of cryptography is old, as demand for privacy is as old as
communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
the character transposition technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}.

The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
computing advances in the 1980s, it for the first time put the power of strong encryption in the hands of ordinary
people. However, the privacy provided by cryptography is a powerful tool, and privacy is a property governments strive
to regulate. The use of encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was
convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason
tried to force testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}.
Government's natural stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of
strong encryption.

@@

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. In 1991, the
U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by law; in
response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to secure email, in order that strong
cryptography be ``made available to the American public before it became illegal to use'' \cite{zimmermann_1996}. In
1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with the goal of ``providing the
public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher
discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
died.

Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
\cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went further, including
``provisions that would have barred the government from creating a mandatory key escrow system and would also have
removed export restrictions on most generally available software and hardware containing encryption''
\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
November of that year, the Clinton administration released an executive order effectively removing export controls on
strong encryption products, along the lines of the \ac{PROCODEAct} \cite{clinton_1996}.

@@

The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
of the September 11 attacks to some justified yet greater investigatory powers \cite{bloss_escalating_2007}. October of
2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
\ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
compel information from digital service providers while precluding any public disclosure of the event
\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.

If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, it was the 2013 \ac{snowden}
revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
promised a technical solution. Under pressure from both the human rights field and from ordinary users demanding
improved digital privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and
\acl{E2E} encryption for \acl{DIM} \cite{treguer_us_2018}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
law enforcement, and cybersecurity, the pursuit of tech professionals. Both sides of the debate have by now acknowledged
that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public security vs.
cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto wars,
terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators jointly
released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it
to the floor, and the unresolved nature of the court case left the legal debate wide open.

\paragraph*{\Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
\cite{may_1994}. The traditional list includes terrorists, pedophiles, drug dealers, and money launderers, though
sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography
\cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on child
pornography is likely a result of the growth of the ``dark web'' and a severe, in-depth 2019 investigation by the
\textit{New York Times} \cite{keller_internet_2019}.

@@

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
like ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020, the
Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to establish a
16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child pornography
\cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely see the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}.

@@

The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.

>>>>

\Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
performed by using a \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. A cipher is a tool or
algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
cipher ensures that only those parties with the correct secret information, the \ac{key}, can perform encryption or
decryption on the text.

@@

There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was discovered
long ago, encryption and decryption are performed with the same key, and both parties must have this key in order to
communicate securely. In \ac{asym-crypto}, which was discovered recently and is also called \ac{public-key-crypto},
encryption and decryption are performed with two paired keys, called the public key and the private key. The innovation
of \ac{public-key-crypto} is that the public key is not secret, and using this technique, two parties can communicate
securely without requiring a previously agreed upon shared secret.

One can also define two major cryptographic applications--securing \acf{DIM} and \acf{DAR}. Each application is a
different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
encryption of the data in transit. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}
and \ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise
any other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an
attacker without detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of
encryption for \ac{DIM}.

\Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
somewhere in computer memory, which often takes place with the assistance of dedicated hardware. \Ac{disk-encryption}
for laptops and mobile devices is an example of encryption for \ac{DAR}.

@@

Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
% FIXME: reading note: paragraph makes sense, but need to read it multiple times for it to sink in. Improve.

@@

Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
systems that directly rely on these technologies include the local and global financial systems; industrial control
systems and critical infrastructure; health information systems; private, government, and military communications; and
the integrity of the internet as a whole.
% FIXME: Jumping rather quickly from primitives to huge systems. Perhaps scale it more smoothly. Connect little
% primitives to higher technologies, higher technologies to underpinnings of network computer systems, and network
% computer systems as critical to these industries.

@@

The section provides a brief overview of encryption from ancient to modern times.

@@

Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
the character transposition technique Julius Caesar used to protect private correspondence
\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}.

The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
rapid increase in availability of strong encryption.
% FIXME: The point is made--these developments made gov'ts uncomfortable (just look at these examples!)--but the line of
% thought is not clear enough to the reader.

@@

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
foundational role of cryptography in security--and the potential of human rights abuses for compromised
privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.

Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
various key escrow proposals \cite{thompson_2015}.
% FIXME: export controls were a hot topic of the time, too. Can't just drop that thread.
In 1996, two pro-encryption bills were introduced in the US Congress.
In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
\cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went further, including
``provisions that would have barred the government from creating a mandatory key escrow system and would also have
removed export restrictions on most generally available software and hardware containing encryption''
\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
November of that year, the Clinton administration released an executive order effectively removing export controls on
strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.

@@

The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
of the September 11 attacks seemed to justify yet greater investigatory powers \cite{bloss_escalating_2007}. October of
2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
\ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
% FIXME: introduce FISA above, in the "had already been expanding" section. Then we have more of a narrative of
% expanding power.
authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
compel information from digital service providers while precluding any public disclosure of the event
\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.

The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, but it was the 2013 \ac{snowden}
revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
promised a technical solution. Benefiting from advancements in underlying technologies and under pressure from both the
human rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing
default device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% FIXME: determine correct usage of -- vs ---
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

\paragraph*{\Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
\cite{may_1994}. The traditional list includes terrorists, pedophiles, drug dealers, and money launderers, though
sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography
\cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on child
pornography is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
York Times} \cite{keller_internet_2019}.

@@

That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
largely see the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
\cite{pfefferkorn_2020}.

@@

The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.
% FIXME: remove this sentence?
